
\begin{acknowledgements}
	I would like to extend my gratitude and appreciation to my advisors Dr. Marc
	Casas and Dr. Eduard Ayguad\'{e}, without their patience and guidance this
	thesis would not have been possible. I would also like to thank Dr.
	Cristiano Malossi and Dr. Panagiotis Hadjidoukas from IBM Z\"{u}rich Lab
	who, during the course of our collaboration, gave me valuable feedbacks and
	insights of my work. As well as for their hospitality during my month-long
	stay in Z\"{u}rich that warmed a lone traveler's heart.

	It has been a long five-year journey. One does not simply navigate through
	those times alone, not to mention the place I call home is ten thousand KM
	away. I want to thank those who were or still are by my side giving me
	support, talking me through things when the getting is tough. My
	appreciation goes to my bestie Kallia, along with Rajiv, Danai, Tomasz and
	David with whom we shared invaluable moments and countless chuckles during
	the good part of my PhD. To Ariel, a close friend that we get along well enough
	to consider share an apartment with. To Ilia and Klaudia who we shared a close
	connection despite the few time we spent. Last but not least, a group of
	people of all ages and from all walks of life yet have enough synergy to
	form a close group that I came across over the final months of my PhD: Alex,
	Asaf, Ettore, John "El Cucho" Osorio, Louis "LeDur" Ledoux, Robin and
	Tamara. A group that offers laughters and stories that makes an otherwise
	stressful ending of a PhD a lot more entertaining.	

	%This thesis has been supported by the Spanish Government (Severo Ochoa grants
	%SEV2015-0493, SEV-2011-00067), by the Spanish Ministry of Science and
	%Innovation (contracts TIN2015-65316-P), by Generalitat de Catalunya
	%(contracts 2014-SGR-1051 and 2014-SGR-1272), by the RoMoL ERC Advanced
	%Grant (GA 321253) and the European HiPEAC Network of Excellence.  This
	%work was also partially performed under the auspices of the U.S.
	%Department of Energy by Lawrence Livermore National Laboratory under
	%Contract DE-AC52-07NA27344 (LLNL-CONF-689878).
\end{acknowledgements}
